---
title: "Lecture 1"
output: html_document
date: "2024-07-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Lecture 1: Regular Typing
These are some notes. Teaching is confusing me. I don't understand what is happening.

제가 공부하고있어요.

- This is bullet 1
- This is bullet 2

1. This is item 1
2. This is item 2

This is **bold** text.
This is *italic* text.

## Lecture 1: R coding

Three backticks: `

```{r}
2+2
5*40
whateveryouwant <- 5*40
```

Some more notes...why do we care about "object assignment operators" or saving things to "objects"?

Because, we can then refer to them and use them **later**.

```{r}
whateveryouwant
```

More sophisticated functions:
- average
- sum

(Aside: Fast commands)
1. PC: ctrl + alt + i
2. Mac: ctrl + option + i

Examples of averaging and summing

- First create 3 numbers and store them in an object
```{r}
threeNumbs <- c(10,-100,50)
threeNumbs
```

- Second take the average with the `mean()` function

```{r}
mean(threeNumbs)
mean(c(10,-100,50))
```

- Or take the median with the `median()` function

```{r}
median(threeNumbs)
```

- Or take the sum with the `sum()` function

```{r}
sum(threeNumbs)
```

Create your own function

```{r}
customMean <- function(x) {
  sum(x) / length(x)
}
```

## Packages

Packages are collections of functions written by someone else.

`tidyverse`

1. Step 1: `install.packages("tidyverse")`
  
  - **DO THIS FROM THE CONSOLE...NOT IN A CODE CHUNK**
  
2. Step 2: `require(tidyverse)`

```{r}
require(tidyverse)
```

## Loading data

We use the `read_rds()` function to load data

```{r}
debt <- read_rds("https://github.com/jbisbee1/ISP_Data_Science_2024/raw/main/data/sc_debt.Rds")

debt
```

